A function--sheaf dictionary for algebraic tori over local fields

### 2014-01-20 shanghai.tex
v2 for arXiv
Replaced Category/iso with \pi_0(Category). What do you think?
Fixed an internal reference

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\QC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\QCb}{{\QC_0}}
\newcommand{\QCf}{{\QC_f}}
\newcommand{\QCiso}[1]{\QC(#1)_{/\textit{iso}}}
\newcommand{\QCbiso}[1]{\QCb(#1)_{/\text{iso}}}
\newcommand{\QCfiso}[1]{\QCf(#1)_{/\text{iso}}}
\newcommand{\trFrob}[1]{t_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

We list some fundamental properties of $f_n$ and $\GN{T}_n$ for further use:
 \begin{enumerate}
  \labitem{(GN.0)}{truncation}
  $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a surjective morphism of group schemes over $\Fq$
  \citelist{\cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276} \cite{stasinski:12a}*{\S 4}}.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a $\QC$-scheme. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$.
  In fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
\end{enumerate}

@@

  Finally, we show that $\trFrob{T,n}$ is functorial.
The first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor,
while the second part shows that trace of Frobenius is a natural transformation
$t_{ - ,n} : F_1 \to F_2$. When further combined with Lemma~\ref{prop:product},
we see that $F_1$ is an additive functor and $t_{ - ,n}: F_1 \to F_2$ is a
natural transformation between additive functors.
\end{proof}

>>>>

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\qcs}[1]{{\mathcal{#1}}}
\newcommand{\gqcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gqcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\qcs{#1}^\dagger}
%% Categories
\newcommand{\QC}{{\mathcal{Q\hskip-0.8pt C}}}
\newcommand{\QCb}{{\QC_0}}
\newcommand{\QCf}{{\QC_f}}
\newcommand{\QCiso}[1]{\pi_0 (\QC(#1))}
\newcommand{\QCbiso}[1]{\pi_0(\QCb(#1))}
\newcommand{\QCfiso}[1]{\pi_0(\QCf(#1))}
\newcommand{\trFrob}[1]{t_{#1}}
%% amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{\bar{g}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gqcs{L} \tight{0}{\boxtimes}{0} \gqcs{L}$}}
\newcommand{\ExE}{\text{$\qcs{E}\tight{0}{\boxtimes}{0}\qcs{E}$}}
\newcommand{\bExE}{\text{$\gqcs{E}\tight{0}{\boxtimes}{0}\gqcs{E}$}}
\newcommand{\AxA}{\text{$A \tight{1}{\times}{1} A$}}
\newcommand{\BxB}{\text{$B \tight{1}{\times}{1} B$}}
\newcommand{\GzxGz}{\text{$G^0 \tight{1}{\times}{1} G^0$}}
%% Margin notes
\newcommand\Clifton[1]{\marginpar{\smaller\smaller CC: #1}}
\newcommand\David[1]{\marginpar{\smaller\smaller DR: #1}}
%% Hyphenation override
\hyphenation{quasi-character}

@@

We list some fundamental properties of $f_n$ and $\GN{T}_n$ for further use:
 \begin{enumerate}
  \labitem{(GN.0)}{truncation}
  $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is a surjective morphism of group schemes over $\Fq$
  \citelist{\cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276} \cite{bertrapelle-gonzales:Greenberg}*{Prop. 7.2}}.
  We remark that $f_n : \GN{T}_{n+1} \to \GN{T}_n$ is not induced from a morphism of Néron models.
  \labitem{(GN.1)}{smooth} 
  $\GN{T}_n$ is a $\QC$-scheme. 
  Note that $\underline{T}$ is a smooth group scheme over $\OK$, by construction. 
  Since smoothness is preserved under base change, 
  $\TT_n \ceq \underline{T} \times_\Spec{\OK} \Spec{\OK/\pK^n}$ 
  is a smooth group scheme over an Artinian ring with perfect residue field. 
  The Greenberg functor preserves smoothness, 
  so $\GN{T}_n$ is a smooth group scheme locally of finite type over $\Fq$. 
  The Greenberg functor takes the geometric identity component of $\TT_n$ 
  to the geometric identity component of $\GN{T}_n$.
  In fact, $\pi_0(\GN{T}_n) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$ and
  this group is finitely generated \cite{xarles:93a}*{2.18}.
  \labitem{(GN.2)}{points} $\GN{T}_n(\Fq) = \TT(\OK/\pK^n)$, canonically.
\end{enumerate}

@@

  Finally, we show that $\trFrob{T,n}$ is functorial.
The first part of Proposition~\ref{prop:pullback} shows that $F_1$ is a functor,
while the second part shows that trace of Frobenius is a natural transformation
$t_{ - ,n} : F_1 \to F_2$. When further combined with Proposition~\ref{prop:product},
we see that $F_1$ is an additive functor and $t_{ - ,n}: F_1 \to F_2$ is a
natural transformation between additive functors.
\end{proof}
